---
title: "Homework 1"
author: "Kevin Phan"
format: html
editor: visual
number-sections: true
---

1.  Generate 1000 random draws from a standard normal distribution and store the values in a vector called x.

    ```{r}
    set.seed(67)
    x <- rnorm(1000)
    ```

    a.  Remove all the values of this vector between 0 and 1. How many values remain in the vector? Compute the mean of the remaining values.

        ```{r}
        length(x[x <= 0 | x >= 1])
        mean(x[x <= 0 | x >= 1])
        ```

    b.  Write a function that computes this mean as a function of n. Run this function once, for n = 100.

        ```{r}
        normal_mean <- function(n) {
          x <- rnorm(n)
          return(mean(x[x <= 0 | x >= 1]))
        }

        normal_mean(100)
        ```

    c.  Using a for loop, run this function 500 times with n = 10. Store these 500 values in a vector and then plot a histogram of these values. Report the largest and smallest values from the 500 runs.

        ```{r}
        y <- numeric(500)
        for (i in 1:500) {
          y[i] <- normal_mean(10)
        }

        hist(y)
        max(y)
        min(y)
        ```

2.  The game of craps is played as follows. First, you roll two six-sided dice; let x be the sum of the dice on the first roll. If x = 7 or 11 you win immediately, if x = 2, 3, or 12 you lose immediately, otherwise you keep rolling until either you get x again, in which case you also win, or until you get a 7, in which case you lose.

    Write a program to simulate a game of craps and estimate the probability that a player wins. You can use the following snippet of code to simulate the roll of two (fair) dice:

    ``` r
    x <- sum(sample(1:6, 2, replace = TRUE))
    ```

    ```{r}
    craps_win <- function() {
      roll_1 <- sum(sample(1:6, 2, replace = TRUE))
      if (roll_1 == 2 | roll_1 == 3 | roll_1 == 12) {
        return(FALSE)
      } else if (roll_1 == 7 | roll_1 == 11) {
        return(TRUE)
      } else {
        next_roll <- sum(sample(1:6, 2, replace = TRUE))
        while(next_roll != roll_1) {
          if (next_roll == 7) {
            return(FALSE)
          } else {
            next_roll <- sum(sample(1:6, 2, replace = TRUE))
          }
        }
        return(TRUE)
      }
    }

    win_pct <- function(n) {
      games <- logical(n)
      for (i in 1:n) {
        games[i] <- craps_win()
      }
      return(sum(games) / n)
    }

    win_pct(100)
    win_pct(1000)
    win_pct(10000)
    win_pct(100000)
    win_pct(1000000)
    ```

    *As the number of craps games played increases, the percentage of craps games won seems to converge to around 49%. Therefore, by the law of large numbers, I estimate the probability of winning to be about 0.49.*

3.  Using a while loop, write a function that counts the number of times a 6-sided die must be rolled until you get a six and returns the number of rolls. Run this function a large number of times to estimate the average number of times a die must be rolled until you get a 6.

    ```{r}
    rolls_until_6 <- function() {
      rolls <- NULL
      roll <- sample(1:6, 1)
      num_rolls <- 1
      rolls <- c(rolls, roll)
      while (roll != 6) {
        roll <- sample(1:6, 1)
        num_rolls <- num_rolls + 1
        rolls <- c(rolls, roll)
      }
      return(list(rolls, num_rolls))
    }

    rolls_until_6()
    ```

4.  
